\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-23 paper.qmd
Rewrite the tour section

A tour is a powerful visualization technique used to explore high-dimensional data by generating a sequence of projections, typically into two dimensions. There are two main types of tours: the Grand Tour and the Guided Tour. The Grand Tour explores the data's shape and global structure by using random projections [@Asimov1985]. In contrast, the Guided Tour focuses on specific patterns by moving towards interesting projections defined by a predefined index function [@article29].

The process begins with a real data matrix $X$ containing $n$ observations in $p$ dimensions. It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases) $A_t$, usually $d$ is $1$ or $2$ dimensions. For each pair of orthonormal bases $A_t$ and $A_{t+1}$, a geodesic path is interpolated to create smooth animation between projections.

In the Grand Tour, new orthonormal bases are randomly chosen to explore the $d$-dimensional subspace. The data is often sphered via principal components to reduce dimensionality. The Guided Tour uses a predefined index function to generate a sequence of 'interesting' projections. The resulting tour continuously visualizes the projected data $Y_t$ = $XA_t$ as it interpolates between successive bases.

While both tours can be used to visualize data, examples often focus on using the Grand Tour to observe global structures. However, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives.

@@

@@

p2_n + p2_n_rm +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

@@

plot_list2_harps <- tSNE_data_harps |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = cluster))+
    geom_point(alpha=0.5) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "cluster"),
         color = guide_legend(nrow = 3)) #+
  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

@@

trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

>>>>

A tour is a powerful visualization technique used to explore the shape and global structure of high-dimensional data by generating a sequence of projections, typically into two dimensions. There are two main types of tours: the grand tour [@Asimov1985] and the guided tour [@article29]. A grand tour involves randomly selecting new orthonormal bases, enabling users to understand the structure by exploring the subspace of d-dimensional projections [@Asimov1985]. In contrast, a guided tour can be employed to generate a sequence of 'interesting' projections based on an index function [@article29].

The process begins with the data matrix $X$. It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases) $P_t$, usually $d$ is one or two dimensions. For each pair of orthonormal bases $P_t$ and $P_{t+1}$, a geodesic path is interpolated to create smooth animation between projections. The resulting tour continuously visualizes the projected data $Y_t$ = $XP_t$ as it interpolates between successive bases.

Furthermore, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives. In our context, use grand tour along with the model to observe how effectively the model captures the underlying structure of the data.

<!--While both tours can be used to visualize data, examples often focus on using the grand tour to observe global structures. However, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives. -->

@@

num_bins_umap_s_curve <- 11
shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                          x = UMAP1, y = UMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)

df_bin_centroids_umap_s_curve <- df_bin_centroids_umap_s_curve |>
  dplyr::filter(std_counts > 0.12)

trimesh_umap_s_curve_rm_low <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
  geom_point(size = 2, colour = "#33a02c") +
  coord_equal()

trimesh_umap_s_curve_rm_low <- trimesh_umap_s_curve_rm_low +
  coord_cartesian(xlim =c(-5, 7.5), ylim = c(-10, 10)) +
  #ggtitle("(a)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  ) #+
  #annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

p2_n + p2_n_rm + trimesh_umap_s_curve_rm_low +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

@@

plot_list2_harps <- tSNE_data_harps |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = cluster))+
    geom_point(alpha=0.5) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "cluster"),
         color = guide_legend(nrow = 3)) +
 annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

### 2024-01-23 paper.tex
Rewrite the tour section

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi

@@

A tour is a powerful visualization technique used to explore
high-dimensional data by generating a sequence of projections, typically
into two dimensions. There are two main types of tours: the Grand Tour
and the Guided Tour. The Grand Tour explores the data's shape and global
structure by using random projections \citep{Asimov1985}. In contrast,
the Guided Tour focuses on specific patterns by moving towards
interesting projections defined by a predefined index function
\citep{article29}.

The process begins with a real data matrix \(X\) containing \(n\)
observations in \(p\) dimensions. It generates a sequence of \(p\) ×
\(d\) orthonormal projection matrices (bases) \(A_t\), usually \(d\) is
\(1\) or \(2\) dimensions. For each pair of orthonormal bases \(A_t\)
and \(A_{t+1}\), a geodesic path is interpolated to create smooth
animation between projections.

In the Grand Tour, new orthonormal bases are randomly chosen to explore
the \(d\)-dimensional subspace. The data is often sphered via principal
components to reduce dimensionality. The Guided Tour uses a predefined
index function to generate a sequence of `interesting' projections. The
resulting tour continuously visualizes the projected data \(Y_t\) =
\(XA_t\) as it interpolates between successive bases.

While both tours can be used to visualize data, examples often focus on
using the Grand Tour to observe global structures. However, software
like \textbf{langevitour} can visualize both types of tours, providing
flexibility for exploring high-dimensional data with various objectives.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, boxrule=0pt, interior hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

A tour is a powerful visualization technique used to explore the shape
and global structure of high-dimensional data by generating a sequence
of projections, typically into two dimensions. There are two main types
of tours: the grand tour \citep{Asimov1985} and the guided tour
\citep{article29}. A grand tour involves randomly selecting new
orthonormal bases, enabling users to understand the structure by
exploring the subspace of d-dimensional projections \citep{Asimov1985}.
In contrast, a guided tour can be employed to generate a sequence of
`interesting' projections based on an index function \citep{article29}.

The process begins with the data matrix \(X\). It generates a sequence
of \(p\) × \(d\) orthonormal projection matrices (bases) \(P_t\),
usually \(d\) is one or two dimensions. For each pair of orthonormal
bases \(P_t\) and \(P_{t+1}\), a geodesic path is interpolated to create
smooth animation between projections. The resulting tour continuously
visualizes the projected data \(Y_t\) = \(XP_t\) as it interpolates
between successive bases.

Furthermore, software like \textbf{langevitour} can visualize both types
of tours, providing flexibility for exploring high-dimensional data with
various objectives. In our context, use grand tour along with the model
to observe how effectively the model captures the underlying structure
of the data.
